Figures

### 2016-08-12 coralmaps/coral_maps.Rmd

if(basename(getwd()) != 'data_explore') setwd('data_explore')
source('data_explore_fxns.R')
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

@@

nrow(spp_depth %>% filter(max_depth > 100))
# [1] 5
### of 562 species, only 5 are listed as occurring below 100 m; none are below 200 m

@@

### 17 species (of 396) have no depth information; one is listed as being dredged from 13 fathoms (~ 25 m)
```

@@

bath_df <-  read_csv(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'), col_types = 'dd') %>%
  filter(weight >= 1) %>%
  mutate(deep = 1) %>%
  select(-weight)

@@

write_csv(coral_cells_iucn, file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv'))

``` {r calc_areas_for_iucn_corals}
### calc areas
### read hcaf to get areas; left join to coral_cells_iucn
### group_by iucn_sid and depth, summarize area (deep and shallow)
### group_by iucn_sid, summarize total area
coral_cells_iucn <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv'))

@@

iucn_coral_areas <- coral_cells_iucn %>%
  left_join(hcaf, by = 'loiczid') %>%
  group_by(iucn_sid, deep) %>%
  summarize(deeparea = sum(oceanarea)) %>%
  group_by(iucn_sid) %>%
  mutate(totarea = sum(deeparea),
         shallow_pct = (totarea - deeparea)/totarea) %>%
  arrange(desc(deep)) %>%
  .[!duplicated(.$iucn_sid), ]

@@

  ### get species list
  coral_cells_iucn_shallow <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv')) %>%
    filter(deep == 0) %>%
    mutate(prop_area = 1)

@@

  write_csv(spp_coralmaps, file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))
}
```

@@

spp_coral_clip <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_areas.csv')) %>%
  select(iucn_sid, shallow_pct, totarea, deeparea) %>%
  mutate(shallow_pct = 100 * shallow_pct)

spp_small_effect <- spp_coral_clip %>%
  filter(shallow_pct > 80) 
### only five with such small effects; one of those did not change at all.

summary(spp_coral_clip$shallow_pct)
  #  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  # 23.80   52.89   55.48   56.33   60.59  100.00

scatterplot1 <- ggplot(spp_coral_clip, aes(x = totalarea, y = deeparea)) +
  geom_point(size = 0.25, color = 'black') +
  geom_abline(color = 'red')
scatterplot1

@@

spp_coral_clip2 <- spp_coral_clip %>%
  select(iucn_sid, shallow = shallow_pct) %>%
  arrange(desc(shallow))

spp_coral_clip2$l_to_h <- c(1:nrow(spp_coral_clip))

spp_coral_clip2 <- spp_coral_clip2 %>%
  mutate(deep = 100 - shallow) %>%
  gather(depth, pct_value, c(shallow, deep)) %>%
  mutate(depth = ifelse(depth == 'shallow', '0 - 200 m', '200 m +'))

break_nums <- seq(0, 100, 20)
coral_area_plot <- ggplot(spp_coral_clip2, aes(x = l_to_h, y = pct_value, fill = depth)) +
  ggtheme_plot +
  # geom_linerange(aes(ymin = kept_ratio, ymax = 100),
  #                position = position_dodge(width = 0.5)) +
  geom_area() +
  scale_fill_manual(values = c('0 - 200 m' = 'grey90', '200 m +' = 'grey10')) +
  geom_hline(yintercept = break_nums, color = 'white', size = .25) +
  scale_y_continuous(limits = c(0, 100), breaks = break_nums,
                     labels = paste0(break_nums, '%')) +
  # scale_x_discrete(labels = element_blank()) +
  # theme(axis.title.x = element_blank()) +
  labs(y = 'IUCN % of coral range',
       x = 'Count of coral species')
print(coral_area_plot)

@@

### recalc areas and overlaps
iucn_coral_fixed <- spp_coralmaps %>%
  gather(dist, perc, c(sm_perc_raw, sm_perc_shallow)) %>%
  gather(area, ratio, c(area_ratio_raw, area_ratio_shallow)) %>%
  filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
  filter(!(str_detect(dist, 'clip') & !(str_detect(area, 'clip')))) %>%
  mutate(method = ifelse(str_detect(dist, 'clip'), '200 m clip', 'original'))

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))

>>>>

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

@@

nrow(spp_depth %>% filter(max_depth > 100))
# [1] 5
### of 562 species, only 5 are listed as occurring below 100 m; none are below 200 m.
### NOTE: Many of these 562 species have no maps in IUCN; they are mapped in AquaMaps though.

@@

### 18 species (of 562) have no depth information; one is listed as being dredged from 13 fathoms (~ 25 m)
```

@@

bath_df <-  read_csv(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'), col_types = 'dd') %>%
  filter(weight >= 1) %>% ### cell must be 100% within the deep polygon to count as deep
  mutate(deep = 1) %>%
  select(-weight)

@@

write_csv(coral_cells_iucn, file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv'))
length(unique(coral_cells_iucn$iucn_sid))
### NOTE: only 473 coral species with IUCN maps
```

``` {r calc_areas_for_iucn_corals}
### calc areas
### read hcaf to get areas; left join to coral_cells_iucn
### group_by iucn_sid and depth, summarize area (deep and shallow)
### group_by iucn_sid, summarize total area
coral_cells_iucn <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv')) %>%
  mutate(deep = ifelse(deep == 1, 'deep', 'shallow'))

@@

iucn_coral_areas <- coral_cells_iucn %>%
  left_join(hcaf, by = 'loiczid') %>%
  group_by(iucn_sid, deep) %>%
  summarize(area = sum(oceanarea)) %>%
  group_by(iucn_sid) %>%
  mutate(totarea = sum(area),
         pct_area = area/totarea) %>%
  filter(deep == 'shallow') %>%
  rename(shallow_pct = pct_area) %>%
  mutate(deep_pct = 1 - shallow_pct)

@@

  ### get species list
  coral_cells_iucn_shallow <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv')) %>%
    filter(deep == 0)

@@

  write_csv(spp_coralmaps, file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))
}

@@

spp_coral_clip <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_areas.csv')) %>%
  select(iucn_sid, shallow_pct, deep_pct, totarea, shallow_area = area) %>%
  mutate(shallow_pct = 100 * shallow_pct,
         deep_pct    = 100 * deep_pct)

spp_small_effect <- spp_coral_clip %>%
  filter(shallow_pct > 80) 
### only four out of 481 with such small effects; one of those did not change at all.

summary(spp_coral_clip$shallow_pct)
  # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  # 22.40   48.20   51.71   52.37   56.68  100.00

scatterplot1 <- ggplot(spp_coral_clip, aes(x = totarea, y = shallow_area)) +
  geom_point(size = 0.25, color = 'black') +
  geom_abline(color = 'red')
scatterplot1

@@

### for area plot, arrange from highest shallow pct to lowest (best to worst)
spp_coral_clip <- spp_coral_clip %>%
  arrange(desc(shallow_pct))

spp_coral_clip$l_to_h <- c(1:nrow(spp_coral_clip))

spp_coral_clip2 <- spp_coral_clip %>%
  gather(depth, pct_value, c(shallow_pct, deep_pct)) %>%
  mutate(depth = ifelse(depth == 'shallow_pct', '0 - 200 m', '200 m +'))

break_nums <- seq(0, 100, 20)
coral_area_plot <- ggplot(spp_coral_clip2, aes(x = l_to_h, y = pct_value, fill = depth)) +
  ggtheme_plot +
  # geom_linerange(aes(ymin = kept_ratio, ymax = 100),
  #                position = position_dodge(width = 0.5)) +
  geom_area() +
  scale_fill_manual(values = c('0 - 200 m' = 'grey90', '200 m +' = 'grey10')) +
  geom_hline(yintercept = break_nums, color = 'white', size = .25) +
  scale_y_continuous(# limits = c(0, 100), ### why is this causing a vertical line?
                     breaks = break_nums,
                     labels = paste0(break_nums, '%')) +
  # scale_x_discrete(labels = element_blank()) +
  # theme(axis.title.x = element_blank()) +
  labs(y = 'IUCN % of coral range',
       x = 'Count of coral species')
print(coral_area_plot)

@@

### recalc areas and overlaps
iucn_coral_fixed <- spp_coralmaps %>%
  gather(dist, perc, c(sm_perc_raw, sm_perc_shallow)) %>%
  gather(area, ratio, c(area_ratio_raw, area_ratio_shallow)) %>%
  filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
  filter(!(str_detect(dist, 'shallow') & !(str_detect(area, 'shallow')))) %>%
  mutate(method = ifelse(str_detect(dist, 'shallow'), '200 m clip', 'original'))

### 2016-08-12 data_explore/data_prep.Rmd

iucn_spp_cells <- iucn_spp_cells_all %>%
  rename(iucn_sid = id_no) %>%
  filter(iucn_sid %in% spp_am_v_iucn$iucn_sid) %>%
  filter(presence != 5)
### length(unique(iucn_spp_cells$iucn_sid)) results in 2279 species found...
### some species with no cells reported.  2279 listed in the am v iucn file.
write_csv(iucn_spp_cells, file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'))

message('Loading AquaMaps species to cell table.  This might take a few minutes.')
am_spp_cells_all <- fread(file.path(dir_data_remote, 'aquamaps/v2015/csv/hcaf_sp_native_trunc.csv'), 
                         showProgress = TRUE)
am_spp_cells <- am_spp_cells_all %>%
  rename(am_sid = speciesid, prob = probability) %>%
  filter(am_sid %in% spp_am_v_iucn$am_sid)
### length(unique(am_spp_cells$am_sid)) results in 2314 species found - great!
write_csv(am_spp_cells, file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'))

@@

  if(!exists('am_spp_cells'))
    am_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'))
  if(!exists('iucn_spp_cells'))
    iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'),
                               col_types = 'cddddc')
  ### species info list - scinames, IUCN species ID, AM species ID, category etc
  spp_list_base <- read_csv(file.path(dir_data, 'spp_am_v_iucn.csv'))

>>>>

iucn_spp_cells <- iucn_spp_cells_all %>%
  rename(iucn_sid = id_no) %>%
  filter(iucn_sid %in% spp_am_v_iucn$iucn_sid) %>%
  filter(presence != 5)
### length(unique(iucn_spp_cells$iucn_sid)) results in 2279 species found...
### some species with no cells reported.  2279 listed in the am v iucn file.
write_csv(iucn_spp_cells, file.path(dir_anx, 'vAM_IUCN/iucn_spp_cells_2015.csv'))

message('Loading AquaMaps species to cell table.  This might take a few minutes.')
am_spp_cells_all <- fread(file.path(dir_data_remote, 'aquamaps/v2015/csv/hcaf_sp_native_trunc.csv'), 
                         showProgress = TRUE)
am_spp_cells <- am_spp_cells_all %>%
  rename(am_sid = speciesid, prob = probability) %>%
  filter(am_sid %in% spp_am_v_iucn$am_sid)
### length(unique(am_spp_cells$am_sid)) results in 2314 species found - great!
write_csv(am_spp_cells, file.path(dir_anx, 'vAM_IUCN/am_spp_cells_2015.csv'))

@@

  if(!exists('am_spp_cells'))
    am_spp_cells <- read_csv(file.path(dir_anx, 'vAM_IUCN/am_spp_cells_2015.csv'))
  if(!exists('iucn_spp_cells'))
    iucn_spp_cells <- read_csv(file.path(dir_anx, 'vAM_IUCN/iucn_spp_cells_2015.csv'),
                               col_types = 'cddddc')
  ### species info list - scinames, IUCN species ID, AM species ID, category etc
  spp_list_base <- read_csv(file.path(dir_data, 'spp_am_v_iucn.csv'))

### 2016-08-12 ms_3_res_disc.Rmd

___Distribution alignment - overlap as proxy for common understanding; use Bayesian argument to show that these areas are high-confidence areas; area ratio as proxy for errors of commission/omission.  While we cannot know which dataset is closer to the "truth" we can make inferences based on expectations of EOO and AOO maps - is AquaMaps AOO?___

_Distribution alignment - Bayesian description of higher confidence_
_Area ratio is defined as the ratio of the smaller range area to the larger range area.  This can provide insights into commission and omission errors: a large difference in areas implies a greater number of errors in one or both datasets.  While we cannot know which dataset is closer to representing the "truth" of each species' range, we can expect that IUCN (representing EOO) will introduce more commission errors, while AquaMaps (representing AOO) is likely to reduce these at the expense of introducing more omission errors._

>>>>

___Distribution alignment - overlap as proxy for common understanding; these areas are high-confidence areas.  Area ratio as proxy for errors of commission/omission.  While we cannot know which dataset is closer to the "truth" we can make inferences based on expectations of EOO and AOO maps - is AquaMaps AOO?___

_Distribution alignment - higher confidence when both datasets predict presence (or absence) in the same area_
_Area ratio is defined as the ratio of the smaller range area to the larger range area.  This can provide insights into commission and omission errors: a large difference in areas implies a greater number of errors in one or both datasets.  While we cannot know which dataset is closer to representing the "truth" of each species' range, we can expect that IUCN (representing EOO) will introduce more commission errors, while AquaMaps (representing AOO) is likely to reduce these at the expense of introducing more omission errors._
